Eval expressions

### 2021-01-28 analysis/expressions.Rmd
Update to normalization

str_op <- c("paste", "paste0", "str_c")

canonic_expr <- function(exp) {
  if (is.call(exp)) {
    function_name <- exp[[1]]
    function_args <- exp[-1]
    res <- map(function_args, canonic_expr)
    # Function_name can be larger than one is a function with a namespace stats::model or another higher-order function
    # For instance, (function(x, y) x + y)(3, 4)
    if (length(function_name) == 1 && as.character(function_name) %in% arith_func && every(res, function(v) {
      v == "NUM"
    })) {
      return("NUM")
    }
    else if (length(function_name) == 1 && as.character(function_name) %in% str_op && every(res, function(v) {
      v == "STR"
    })) {
      return("STR")
    }
    else if (length(function_name) == 1 && function_name == "::") {
      str_c(as.character(function_args[[1]]), "::", as.character(function_args[[2]]))
    }
    else {
      function_name <- if (length(function_name) == 1) {
        as.character(function_name)
      }
      else {
        canonic_expr(function_name)
      }
      return(str_c(str_c(function_name, collapse = ", "), "(", str_c(res, collapse = ", "), ")"))
    }
  }
  else if (is.symbol(exp)) {
    return("VAR")
  }
  else if (is.expression(exp)) {
    return(str_c(map(exp, canonic_expr), collapse = ", "))
  }
  else if (typeof(exp) %in% c("integer", "double", "complex")) {
    return("NUM")
  }
  else if(typeof(exp) == "logical") {
    return("BOOL")
  }
  else if(typeof(exp) == "character") {
    return("STR")
  }
  else {
    return(deparse1(exp))
  }
}

canonic_expr_str <- function(exp) {
  ast <- NA
  try(ast <- parse(text = exp)[[1]], silent = TRUE)
  # some expr_resolved have been truncated so we mark them as FALSE (even though they could be true)
  if (is.symbol(ast) || is.language(ast) || length(ast) > 1 || !is.na(ast)) {
    return(canonic_expr(ast))
  }
  else {
    return("CANONIZATION ERROR")
  }
}
```

@@

```{r}
res_exprs <- E %>%
  select(expr_resolved) %>%
  unique()

@@

>>>>

```{r}
str_op <- c("paste", "paste0", "str_c")
comp_op <- c("<", ">", "<=", ">=", "==", "!=")
bool_op <- c("&", "&&", "|", "||", "!")

canonic_expr <- function(exp, with.names = FALSE) {
    if (is.call(exp)) {
        function_name <- exp[[1]]
        function_args <- exp[-1]
        res <- map(function_args, canonic_expr, with.names) # TODO: rather directly compute canonic_expr on function_args?
        # Function_name can be larger than one is a function with a namespace stats::model or another higher-order function
        # For instance, (function(x, y) x + y)(3, 4)
        if (length(function_name) == 1 && as.character(function_name) == "(") {
            return(res[[1]]) # (a) => (a)()
        }
        else if (length(function_name) == 1 && as.character(function_name) %in% arith_func && every(res, function(v) {
            v == "NUM"
        })) {
            return("NUM")
        }
        else if (length(function_name) == 1 && as.character(function_name) %in% str_op && every(res, function(v) {
            v == "STR"
        })) {
            return("STR")
        }
        else if (length(function_name) == 1 && as.character(function_name) %in% bool_op && every(res, function(v) {
            v == "BOOL"
        })) {
            return("BOOL")
        }
        else if (length(function_name) == 1 && as.character(function_name) %in% comp_op && every(res, function(v) {
            v == "NUM"
        })) {
            return("BOOL")
        }
        else if (length(function_name) == 1 && as.character(function_name) %in% c("c", "list") && length(res) > 1 && n_distinct(res) == 1 && res[[1]] %in% c("BOOL", "NUM", "STR", "VAR")) {
            str_c(as.character(function_name), "(", res[[1]], ")")
        }
        else if (length(function_name) == 1 && function_name == "::") { # we drop the namespace
            as.character(function_args[[2]])
        }
        else {
            function_name <- if (length(function_name) == 1) {
                func_name <- as.character(function_name)
                function_name <- if (func_name %in% arith_func) {
                    "OP"
                }
                else if (func_name %in% comp_op) {
                    "COMP"
                }
                else if (func_name %in% bool_op) {
                    "LOGI"
                }
                else {
                    func_name
                }
            }
            else {
                canonic_expr(function_name, with.names)
            }

            # It might have been transformed by canonic_expr in the recursive call above
            # it should always have length one
            res <- if (length(function_name) == 1) {
                func_name <- as.character(function_name)
                # VAR is absorbing: if there is at least one, we propagate to all of them
                res <- if (func_name %in% c("COMP", "LOGI", "OP") && "VAR" %in% res) {
                    rep.int("VAR", length(res))
                }
                else if (func_name == "model.frame") { # special case
                    # We keep the 1st two arguments (usually formula and data)
                    # and subset if it is there
                    subset_arg <- if ("subset" %in% names(res)) {
                        subset_s <- as.character(res["subset"])
                        res["subset"] <- NULL
                        str_c("subset = ", subset_s)
                    }
                    else {
                        "subset = NULL"
                    }
                    first_args <- if(length(res) > 1) {
                          res[1:2]
                    }
                    else {
                      c(res[[1]], "NULL")
                    }
                  c(first_args, subset_arg)
                }
                else {
                    res
                }
            }

            fun_args <- if(with.names) {
              str_c(names(res), if_else(names(res) == "", "",  " = "), res, collapse = ", ")
            } 
            else
            {
              str_c(res, collapse = ", ")
            }

            return(str_c(str_c(function_name, collapse = ", "), "(", fun_args, ")"))
        }
    }
    else if (is.symbol(exp)) {
        return("VAR")
    }
    else if (is.expression(exp)) {
        return(str_c(map(exp, canonic_expr, with.names), collapse = ", "))
    }
    else if (typeof(exp) %in% c("integer", "double", "complex")) {
        return("NUM")
    }
    else if (typeof(exp) == "logical") {
        return("BOOL")
    }
    else if (typeof(exp) == "character") {
        return("STR")
    }
    else {
        return(deparse1(exp))
    }
}

canonic_expr_str <- function(exp, with.names = FALSE) {
    ast <- NA
    try(ast <- parse(text = exp)[[1]], silent = TRUE)
    # some expr_resolved have been truncated so we mark them as FALSE (even though they could be true)
    if (is.symbol(ast) || is.language(ast) || length(ast) > 1 || !is.na(ast)) {
        return(canonic_expr(ast, with.names))
    }
    else {
        return("NORMALIZATION ERROR")
    }
}
```

@@

```{r}
res_exprs_uniq <- E %>%
  select(expr_resolved) %>%
  unique()

@@

res_exprs <- res_exprs %>% mutate(expr_canonic_names = map_chr(expr_resolved, canonic_expr_str), with.names = TRUE)
